---
title: levels() not working in 2nd Edition of Getting Started
author: Owen Petchey
date: '2022-09-10'
image:
  caption: ''
  focal_point: ''
---

Sometimes, in the second edition, we use the `levels` function to get the unique levels of a variable. For example on page 133 we do `levels(growth.moo$diet)` to get the unique levels of the `diet` variable. Today, this does not work. Below I explain why and how to fix it. Short version is use `unique` instead of `levels` or convert the variables to factors.

## Prepare

We will use the `mutate` function from the `dplyr` package, so please ensure you have that package installed.

## Import the data

In the next line of code I import the data from github, rather than a local copy. This saves us having to deal with local location of the data file. I would normally work with a local copy, however.

```{r}
growth.moo <- read.csv(url("https://raw.githubusercontent.com/r4all/datasets/master/growth.csv"))
```

## Using `unique` rather than `levels`

Looking at the structure of the data in R we see:

```{r}
str(growth.moo)

```

Supplement and diet are both `chr` (character) type variables.

Hence the `levels` function doesn't give us the levels. Instead, we get `NULL`:

```{r}
levels(growth.moo$supplement)
levels(growth.moo$diet)
```

So, instead use `unique`:

```{r}
unique(growth.moo$supplement)
unique(growth.moo$diet)
```

Awesomeness!

## Converting to a factor

Another option is to convert the `chr` type variables to be factor type variables. There are many ways to achieve this, here are two.

If we want to convert to factors all the `chr` variables in our data, then we can use the `type.convert` function with the argument `as.is = FALSE`. Making this `FALSE` tells the `type.convert` function to *not* keep character type variables *as they are*, but rather to convert them to factors. 

```{r}
growth.moo.factors1 <- type.convert(growth.moo, as.is = FALSE)
str(growth.moo.factors1)
```

Great! What were character type variables are now factors.

By the way, the default since R 4.0.0 is `as.is = TRUE` which can be understood as keep variables as they are--do not convert them to factors. We wrote the second edition before 4.0.0, and this is why `levels` worked when we wrote the second edition, but does not work now.

Another way is to individually convert each variable, for example:

```{r}
growth.moo.factors2  <- dplyr::mutate(growth.moo,
                                      supplement = as.factor(supplement),
                                      diet = as.factor(diet))
str(growth.moo.factors2)
```

Awesomeness 2!

Thanks for reading. Have a nice day!
